---
title: "exam2"
author: "Alex Rigney"
date: "6/26/2020"
output: word_document
---
Importing the data.
```{r, message=FALSE}
library(rio)
inequality_data <- import("inequality.xlsx")
```

Getting to know the data.
```{r}
str(inequality_data)
head(inequality_data)
tail(inequality_data)
mean(inequality_data$year)
```

The data appears to be cross-sectional, as each observation is only sampled once.

Looking at Sweden's Gini score
```{r}
se_gini <- subset(inequality_data, iso2c == "SE", select = inequality_gini)
print(se_gini)

```

Looking at Denmark's score
```{r}
dk_gini <- subset(inequality_data, iso2c == "DK", select = inequality_gini)
print(dk_gini)
```

Looking at Brazil's Gini score
```{r}
br_gini <- subset(inequality_data, iso2c == "BR", select = inequality_gini)
print(br_gini)
```

It is better to have lower Gini Coefficient scores. That means your country's income is distributed more equally. 

Looking at the head of the data.
```{r}
head(inequality_data)
```

Creating a function to remove accents from the country variable. 
```{r}
# define the function
accents.remove <- function(s) {
# 1 character substitutions
old <- "ú"
new <- "u"
s1 <- chartr(old, new, s)

for(i in seq_along(old)) s1 <- gsub(old[i], new[i], s1, fixed = TRUE)
s1
}
# finish the accent FIX
inequality_data$country = accents.remove(inequality_data$country)

head(inequality_data)

```

Sorting the data by Gini score, from lowest to highest.
```{r}
inequality_data <- inequality_data[order(inequality_data$inequality_gini),]

head(inequality_data)
```

Checking the mean of the Gini Score.
```{r}
gini_mean <- mean(inequality_data$inequality_gini, na.rm = TRUE)
gini_mean
```

Trying and failing to create a high and low inequality dummy variable.
```{r}
inequality_data$low_inequality=NA
inequality_data$high_inequality=NA

ifelse(test = inequality_data$inequality_gini > 36.81375, yes = inequality_data$high_inequality <- 1, no = inequality_data$low_inequality <- 1)
```
I've spent 30 minutes on this, I give up


Printing three names using a for loop.
```{r}
names <- c("The World Bank", "the African Development Bank", "the Bill and Melinda Gates Foundation")
for(n in names)
  print(n)
```




I chose the literacy rate for adults over the age of 15. I would assume that nations with a low literacy rate would tend to have higher inequality.

Importing the data and changing the name to something sensable.
```{r}
library(WDI)
lit_data = WDI(country = "all",
indicator = c("SE.ADT.LITR.ZS"), # indicator from web
start = 2015, end = 2015, extra = FALSE, cache = NULL)

library(data.table)
setnames(lit_data, "SE.ADT.LITR.ZS", "lit_rate")
```



Merging the two data frames into one, left side join.
```{r}
library(tidyverse)

merged_df <- left_join(x=inequality_data,
                       y=lit_data,
                       by = c("iso2c", "country", "year"))

str(merged_df)
names(merged_df)
head(merged_df)
```


Dropping all the countries with NAs in either the gini scores or the literacy rate.
```{r}
merged_df <-
merged_df %>%
  filter(!(inequality_gini=="NA")) %>% 
  filter(!(lit_rate=="NA"))

```



Creating a data frame of only those countries with gini scores above 30.
```{r}
data_greater_30 <-
merged_df %>%
  filter(inequality_gini > 30)
data_greater_30
```


summing up the number of countries that have "ai" in their name.
```{r}
sum(grep("ai", data_greater_30))
```

Summing up the gini scores, or at least trying to.
```{r}
sapply(data_greater_30$inequality_gini, FUN = sum)
```


Labeling and exporting the data frame so future researchers will be able to continue my great work.
```{r}
library(labelled)
var_label(data_greater_30) <- list(
      `iso2c` = "boop beep boop",           
      `country` = "different thing",
      `inequality_gini` = "another thing",
      `year` = "How many are there?",
      `low_inequality` = "it can't go on any longer",
      `high_inequality` = "but it is!",
      `lit_rate` = "thanks for grading the test!")

# save the dataset in Stata format with the labels
export(data_greater_30, file = "final_data.dta")
```


https://github.com/adrr1/rwork

Thank you!